Introducing the JuliaHealth Observational Health Research Subecosystem

### 2023-12-31 paper.tex
Wrapped up some more todos

% TODO::MEDIUM Maybe move this into the tutorial code
\begin{listing}[!ht]
\begin{minted}[breaklines,escapeinside=||,mathescape=true, numbersep=3pt, gobble=2, frame=lines, fontsize=\scriptsize, framesep=2mm]{julia}
import Base:
  Fix2

import OMOPCDMCohortCreator:
  GetPatientRace,
  GetPatientGender,
  GetPatientAgeGroup

CGetPatientRace = Fix2(GetPatientRace, c)
CGetPatientGender = Fix2(GetPatientGender, c)
CGetPatientAgeGroup = Fix2(GetPatientAgeGroup, c)
\end{minted}
% TODO::EASY Add a caption
\caption{}
% TODO::EASY Add a label
\label{listing:}
\end{listing}

% TODO::EASY Add this to previous code caption
NOTE: The "C" prefix for each of our curried functions denotes that it is curried.

@@

% TODO::MEDIUM Do we even need chaining here or could we use direct composition of functions?
\begin{listing}[!ht]
\begin{minted}[breaklines,escapeinside=||,mathescape=true, numbersep=3pt, gobble=2, frame=lines, fontsize=\scriptsize, framesep=2mm]{julia}
import Chain:
  @chain

CofactorPatients(patients) = @chain patients begin
    CGetPatientRace_fixed
    CGetPatientGender_fixed
    CGetPatientAgeGroup_fixed
end
\end{minted}
% TODO::EASY Add a caption
\caption{}
% TODO::EASY Add a label
\label{listing:}
\end{listing}

% TODO::MEDIUM Can this be simplified?
\begin{listing}[!ht]
\begin{minted}[breaklines,escapeinside=||,mathescape=true, numbersep=3pt, gobble=2, frame=lines, fontsize=\scriptsize, framesep=2mm]{julia}
CharacterizePatients(patients) = @chain patients begin
    _[:, df.Not(:person_id)]
    df.groupby(_, names(_))
    df.combine(_, df.nrow => :count)
    occ.ExecuteAudit
end
\end{minted}
% TODO::EASY Add a caption
\caption{}
% TODO::EASY Add a label
\label{listing:}
\end{listing}

This gives one the ability to simply reason about an analysis in a very straightforward and readable manner:

\begin{listing}[!ht]
\begin{minted}[breaklines,escapeinside=||,mathescape=true, numbersep=3pt, gobble=2, frame=lines, fontsize=\scriptsize, framesep=2mm]{julia}
RunStudy(patients) = @chain patients begin
    CofactorPatients
    CharacterizePatients
end
\end{minted}
% TODO::EASY Add a caption
\caption{}
% TODO::EASY Add a label
\label{listing:}
\end{listing}

@@

vcat(result...) |> audit_patient_groups
\end{minted}
% TODO::EASY Add a caption
\caption{}
% TODO::EASY Add a label
\label{listing:}
\end{listing}

@@

\begin{listing}[!ht]
\begin{minted}[breaklines,escapeinside=||,mathescape=true, numbersep=3pt, gobble=2, frame=lines, fontsize=\scriptsize, framesep=2mm]{julia}
  GetPatientAgeGroup([1]; age_groupings = [[0, 40], [41, 80]])
\end{minted}
% TODO::EASY Add a caption
\caption{}
% TODO::EASY Add a label
\label{listing:}
\end{listing}

\begin{listing}[!ht]
\begin{minted}[breaklines,escapeinside=||,mathescape=true, numbersep=3pt, gobble=2, frame=lines, fontsize=\scriptsize, framesep=2mm]{julia}
  SELECT
    "PERSON_2"."person_id",
    (CASE 
      WHEN ("PERSON_2"."age" < 41) 
        THEN '0 - 40' 
      WHEN ("PERSON_2"."age" < 81) 
        THEN '41 - 80' END) AS "age_group"
  FROM (
    SELECT
      "PERSON_1"."person_id",
      (2023 - "PERSON_1"."year_of_birth") AS "age"
    FROM "PERSON" AS "PERSON_1"
    WHERE ("PERSON_1"."person_id" IN (1))
  ) AS "PERSON_2"
\end{minted}
% TODO::EASY Add a caption
\caption{}
% TODO::EASY Add a label
\label{listing:}
\end{listing}

@@

\section{References}

% TODO::HARD Review and probably remove appendix

% APPENDIX
\appendix

\section{JSON Serialization of Computable Phenotype Definition}\label{appendix:json}

An example of the JSON serialization of computable phenotype definitions can be found within the paper repository here:

In particular, this example seeks to identify individuals with Type 2 Diabetes Mellitus within a given OMOP CDM.

This was generated using ATLAS Version 2.12.1 as well as 
WebAPI Version 2.12.1.
A generic definition of the schema that was created for ATLAS 2.7.1 can be explored here: https://github.com/OHDSIBr/ATLAS-JSON-Schema/tree/2.7.4

\section{PostgreSQL Serialization of Computable Phenotype Definition}\label{appendix:postgresql}

An example of a PostgreSQL SQL serialization of computable phenotype definitions can be found within the paper repository here:

In particular, this example seeks to identify individuals with Type 2 Diabetes Mellitus within a given OMOP CDM.

This was generated using ATLAS Version 2.12.1 as well as 
WebAPI Version 2.12.1.

>>>>

@@

\begin{listing}[!ht]
\begin{minted}[breaklines,escapeinside=||,mathescape=true, numbersep=3pt, gobble=2, frame=lines, fontsize=\scriptsize, framesep=2mm]{julia}
FGetPatientRace = Fix2(occ.GetPatientRace, conn)
FGetPatientGender = Fix2(occ.GetPatientGender, conn)
FGetPatientAgeGroup = Fix2(occ.GetPatientAgeGroup, conn)

CofactorPatients(patients) = @chain patients begin
    FGetPatientRace
    FGetPatientGender
    FGetPatientAgeGroup
end

CharacterizePatients(patients) = @chain patients begin
    _[:, df.Not(:person_id)]
    df.groupby(_, names(_))
    df.combine(_, df.nrow => :count)
    occ.ExecuteAudit
end

RunStudy(patients) = @chain patients begin
    CofactorPatients
    CharacterizePatients
end
\end{minted}
\caption{\textbf{Modularizing Study Code.} Using Julia's partial application abilities, connection objects can be fixed allowing study code to be readily modularized. Using the package, Chain.jl, occ and processing functions can be composed together to create modular study code.}
\label{listing:modular_study}
\end{listing}

@@

vcat(result...) |> audit_patient_groups
\end{minted}
\caption{\textbf{Distributed Computation of a Study.} With occ, study code can be easily abstracted and modularized. Using Julia's native Distributed package, each module of a study can then be parallelized over patient cohorts.}
\label{listing:study_parallelization}
\end{listing}

@@

\begin{listing}[!ht]
\begin{minted}[breaklines,escapeinside=||,mathescape=true, numbersep=3pt, gobble=2, frame=topline, fontsize=\scriptsize, framesep=2mm]{julia}
# Julia Input
occ.GetPatientAgeGroup([1]; age_groupings = [
                                              [0, 40], 
                                              [41, 80]
                                            ])

\end{minted}

\begin{minted}[breaklines,escapeinside=||,mathescape=true, numbersep=3pt, gobble=2, frame=bottomline, fontsize=\scriptsize, framesep=2mm]{sql}
# SQL Output
SELECT
  "PERSON_2"."person_id",
  (CASE 
    WHEN ("PERSON_2"."age" < 41) 
      THEN '0 - 40' 
    WHEN ("PERSON_2"."age" < 81) 
      THEN '41 - 80' END) AS "age_group"
FROM (
  SELECT
    "PERSON_1"."person_id",
    (2023 - "PERSON_1"."year_of_birth") AS "age"
  FROM "PERSON" AS "PERSON_1"
  WHERE ("PERSON_1"."person_id" IN (1))
) AS "PERSON_2"
\end{minted}
\caption{\textbf{Producing SQL from Julia Expression.} When occ functions are not passed a connection object, they can produce SQL representing the underyling query the Julia expression is executing.}
\label{listing:julia_sql}
\end{listing}

@@
